Comparison of Terrain Following and Cut Cell Grids using a Non-Hydrostatic Model

### 2016-01-05 src/tf-cutcell-comparison/responses2.tex
Remove detailed discussion of horizontal advection
And update advection statistics table for slanted cell grids.

Sections 4a and 4b – Section 4a presents results for the Sch\"{a}r advection test case for various vertical coordinates and advection schemes; there is little new here that is not already covered in the \citet{schaer2002} paper. The advection tests in section 4b provide an interesting counter example to this test. I would recommend switching the emphasis in these two sections, presenting the advected tracer contours in a new Fig. 3 for the terrain-following advection, and just including results from the Sch\"{a}r case in table 1 for comparison purposes.
\end{quotation}
\begin{quotation}
 Lines 131-147 on cut-cell grid generation - It's still not clear how or why a cell vertex  is moved laterally in constructing the cut-cell mesh. Taking Fig. 2 as an example, it  seems that following the authors' directions to move all vertices lying underneath the  orography on the uniform grid up to the surface would produce four grid cells as  shown in the figure below rather than the three cells shown in Fig. 2. This four-cell structure appears to be consistent with the generation of thin cells as the authors  point out in Fig. 5c. If lateral displacement of a vertex is allowed, how is it decided  whether it should be moved and how is it decided where along the surface it should  be located (i.e. the location of point v in Fig. 2)?
\end{quotation}
We have found an alternative method of generating meshes that does not require the use of the \texttt{snappyHexMesh} utility.  This new method does not move vertices laterally and is more straightforward to describe.  This new method creates meshes that retain the long, thin cells from the old method.  The new method is described in section 2 and figure 2.  All test results on cut cell grids have been updated.

@@

\begin{quotation}
 Lines 265-267 and table 1 – The authors are unnecessarily complicating their tests  by comparing results from the cut-cell grid and a regular grid. These grids should be identical for this test and merit no further comparative consideration. There is no need to simulate with both grids, show their identical behavior in table 1, and then  state that the same behavior ``is to be expected.''
\end{quotation}
\TODO{agreed, although the comparison with the fourth order scheme in table 1 is perhaps useful?}

@@

\begin{quotation}
 Lines 364-368 – The authors did not need to run these resting atmosphere 
 simulations to reach the conclusion stated in this paragraph. As stated in my first 
 review, this test case does not challenge the numerics on a cut-cell grid since the horizontal pressure-gradient errors will obviously be reduced in the vicinity of the 
 low level inversion where the coordinates surfaces are perfectly horizontal. A better 
 test case might be to specify an inversion layer that intersects the terrain, since on 
 the cut-cell grid, it's only just adjacent to the terrain where pressure gradient 
 calculations between cell centers are not horizontal.   
\end{quotation}
\TODO{good idea.  zaengl2012 did something similar by raising the mountain to 4km, but we think that simply moving the inversion layer is preferable because the mesh remains constant.  Should we abandon the klemp2011 test entirely or just put less emphasis on its results?}

>>>>

Sections 4a and 4b -- Section 4a presents results for the Sch\"{a}r advection test case for various vertical coordinates and advection schemes; there is little new here that is not already covered in the \citet{schaer2002} paper. The advection tests in section 4b provide an interesting counter example to this test. I would recommend switching the emphasis in these two sections, presenting the advected tracer contours in a new Fig. 3 for the terrain-following advection, and just including results from the Sch\"{a}r case in table 1 for comparison purposes.
\end{quotation}
\TODO{figure 3 will change to show some TF advection results}.  The discussion of the horizontal advection results starting at line \TODO{xxx} has been rewritten more succinctly.

\begin{quotation}
 Lines 131-147 on cut-cell grid generation - It's still not clear how or why a cell vertex  is moved laterally in constructing the cut-cell mesh. Taking Fig. 2 as an example, it  seems that following the authors' directions to move all vertices lying underneath the  orography on the uniform grid up to the surface would produce four grid cells as  shown in the figure below rather than the three cells shown in Fig. 2. This four-cell structure appears to be consistent with the generation of thin cells as the authors  point out in Fig. 5c. If lateral displacement of a vertex is allowed, how is it decided  whether it should be moved and how is it decided where along the surface it should  be located (i.e. the location of point v in Fig. 2)?
\end{quotation}
We have found an alternative method of generating meshes that does not require the use of the \texttt{snappyHexMesh} utility.  This new method does not move vertices laterally and is more straightforward to describe.  This new method creates meshes that retain the long, thin cells from the old method.  The new method is described in section 2 and figure 2.  All test results on cut cell grids have been updated.

@@

\begin{quotation}
 Lines 265-267 and table 1 – The authors are unnecessarily complicating their tests  by comparing results from the cut-cell grid and a regular grid. These grids should be identical for this test and merit no further comparative consideration. There is no need to simulate with both grids, show their identical behavior in table 1, and then  state that the same behavior ``is to be expected.''
\end{quotation}
\TODO{agreed, although the comparison with the fourth order scheme in table 1 is perhaps useful?  I have not removed it yet.}

@@

\begin{quotation}
 Lines 364-368 – The authors did not need to run these resting atmosphere 
 simulations to reach the conclusion stated in this paragraph. As stated in my first 
 review, this test case does not challenge the numerics on a cut-cell grid since the horizontal pressure-gradient errors will obviously be reduced in the vicinity of the 
 low level inversion where the coordinates surfaces are perfectly horizontal. A better 
 test case might be to specify an inversion layer that intersects the terrain, since on 
 the cut-cell grid, it's only just adjacent to the terrain where pressure gradient 
 calculations between cell centers are not horizontal.   
\end{quotation}
\TODO{good idea.  Should we abandon the klemp2011 test entirely or just put less emphasis on its results?}

### 2016-01-05 src/tf-cutcell-comparison/tf-cutcell-comparison.tex
Remove detailed discussion of horizontal advection
And update advection statistics table for slanted cell grids.

Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
The results are compared on the BTF grid for the centred linear scheme (\ref{fig:advection-tracer}a) and the fourth order scheme from \citet{schaer2002} (\ref{fig:advection-tracer}c), and the upwind-biased cubic scheme on the cut cell grid (\ref{fig:advection-tracer}b), and BTF grid (\ref{fig:advection-tracer}d).  Tracer errors at \(t = \SI{10000}{\second}\) are shown for the fourth order and upwind-biased cubic schemes in figures \ref{fig:advection-tracer}e and \ref{fig:advection-tracer}f respectively.

By \(t = \SI{10000}{\second}\), the tracer suffers from distortion on the BTF grid using the centred linear scheme and some artifacts remain about the mountain peak.  The results are visually identical to those using the modified Sch\"{a}r model with centred linear differencing.
The tracer has spread vertically due to increased numerical errors when the tracer is transported between layers.  Distortions are reduced by using the fourth order scheme from \citet{schaer2002} (figure~\ref{fig:advection-tracer}c), but dispersion errors remain with grid-scale oscillations that travel in the opposite direction to the wind (figure~\ref{fig:advection-tracer}e).  The results from the centred linear and fourth order schemes are slightly worse than the respective results from \citet{schaer2002} (their figure 6a and figure 8) because of the difference in the discretisation of the streamfunction, $\Psi$.

Using the upwind-biased cubic scheme, tracer magnitude and shape are well-preserved on all grids, both above the mountain at \(t = \SI{5000}{\second}\) and past the mountain at \(t = \SI{10000}{\second}\).  In this test, advection is most accurate on the cut cell grid (figure~\ref{fig:advection-tracer}b) and regular grid (not shown).  As found by \citet{good2014}, the result is the same on both grids.  This is to be expected since the wind is zero in the region of the ground and flow aloft is aligned with the grids.  On the BTF grid, the tracer is less distorted by the cubic upwind-biased scheme (figure~\ref{fig:advection-tracer}d) compared to the centred linear scheme (figure~\ref{fig:advection-tracer}a) or fourth order scheme (figure~\ref{fig:advection-tracer}c).

@@

The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.  Error is largest on the BTF grid with \(\ell_2 = \num{0.112}\) but is significantly reduced on the SLEVE grid with \(\ell_2 = \num{0.0146}\).  The error is approximately halved by changing from the SLEVE grid to the cut cell grid.
Tracer minima and maxima for the centred linear and fourth order schemes are lower than those presented by \citet{schaer2002} because no interpolation is used to calculate the streamfunction.

@@

The spatial domain, mountain profile, initial tracer profile and discretisation are the same as those in the horizontal tracer advection test.  The velocity field is defined using a streamfunction, $\Psi$, so that the discrete velocity field is non-divergent and follows the BTF coordinate surfaces given by equation~(\ref{eqn:btf}) such that
\begin{equation}
	\Psi(x,z) = -u_0 H \frac{z - h}{H - h} \label{eqn:streamfunc-btf}
\end{equation}
where $u_0 = \SI{10}{\meter\per\second}$, which is the horizontal wind speed where $h(x) = 0$.
The horizontal and vertical components of velocity, $u$ and $w$, are then given by
\begin{align}
	u &= -\frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = \frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{H - z}{\left( H - h \right)^2} \label{eqn:uw-btf} \\
	\frac{\mathrm{d} h}{\mathrm{d} x} &= - h_0 \left[ 
		\beta \cos^2 \left( \alpha x \right) \sin \left( 2 \beta x \right) +
		\alpha \cos^2 \left( \beta x \right) \sin \left( 2 \alpha x \right)
	\right]
\end{align}
Unlike the horizontal advection test, flow extends from the top of the domain all the way to the ground.  The discrete velocity field is calculated using the streamfunction in the same way as the horizontal advection test.

@@

$\ell_2$ errors and tracer extrema for this test are compared with the horizontal advection results in table~\ref{tab:advection}.  In the terrain following velocity field, tracer accuracy is greatest on the BTF grid.  Using the cubic upwind-biased scheme, errors are about ten times larger on the SLEVE and cut cell grids compared to the BTF grid.

@@

\begin{table*}
	\caption{Minimum and maximum tracer densities (\si{\kilogram\per\meter\cubed}) and \(\ell_2\) error norms, defined by equation~(\ref{eqn:l2-error}), at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.  For the horizontal advection test, \(\ell_2\) error norms, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}
\label{tab:advection}
%
\centering
\footnotesize
\begin{tabular}{l l l l S S S S}
\hline\hline
                  &                     &                  & Analytic & {BTF}   & {SLEVE}         & {Cut cell}     & {No terrain} \\
\hline
Horizontal        & Centred linear      & \(\ell_2\) error & 0        & 0.284   & 0.0316          & 999         & 0.0304      \\
                  &                     & min              & 0        & -0.275  & -0.0252         & 999        & -0.0251     \\
                  &                     & max              & 1        & 0.925   & 0.985           & 999          & 0.985       \\
                  & Fourth order        & \(\ell_2\) error & 0        & 0.0938  & 0.00244         & {---}          & 0.00234     \\
                  &                     & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
                  &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 999        & 0.00784     \\
                  &                     & min              & 0        & -0.0464 & -0.0106         & 999      & -0.00674    \\
                  &                     & max              & 1        & 0.922   & 0.982           & 999          & 0.983       \\
\hline
Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0341  & 0.235           & 999     & {---}        \\
	          &                     & min              & 0        & -0.0245 & -0.120          & 999     & {---}        \\
		  &                     & max              & 1        & 0.985   & 0.950           & 999     & {---}        \\
		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0209  & 0.162           & 999          & {---}        \\
                  &                     & min              & 0        & -0.0110 & -0.0263         & 999        & {---}        \\
                  &                     & max              & 1        & 0.983   & 0.865           & 999          & {---}        \\
\hline
\end{tabular}
\end{table*}

>>>>

@@

The results of the cubic upwind-biased scheme on TF and regular grids are comparable with those for the fourth-order centred scheme from \citet{schaer2002}.  Error is largest on the BTF grid with \(\ell_2 = \num{0.112}\) but is significantly reduced on the SLEVE grid with \(\ell_2 = \num{0.0146}\).
Advection is most accurate on the cut cell and regular grids, with \(\ell_2\) approximately half of that on the SLEVE grid.
Tracer minima and maxima for the centred linear and fourth order schemes are lower than those presented by \citet{schaer2002} because no interpolation is used to calculate the streamfunction.

@@

The spatial domain, mountain profile, initial tracer profile and discretisation are the same as those in the horizontal tracer advection test, except for the timestep \(\Delta t = \SI{20}{\second}\).  The velocity field is defined using a streamfunction, $\Psi$, so that the discrete velocity field is non-divergent and follows the BTF coordinate surfaces given by equation~(\ref{eqn:btf}) such that
\begin{equation}
	\Psi(x,z) = -u_0 H \frac{z - h}{H - h} \label{eqn:streamfunc-btf}
\end{equation}
where $u_0 = \SI{10}{\meter\per\second}$, which is the horizontal wind speed where $h(x) = 0$.
The horizontal and vertical components of velocity, $u$ and $w$, are then given by
\begin{align}
	u &= -\frac{\partial \Psi}{\partial z} = u_0 \frac{H}{H - h}, \quad w = \frac{\partial \Psi}{\partial x} = u_0 H \frac{\mathrm{d} h}{\mathrm{d} x} \frac{H - z}{\left( H - h \right)^2} \label{eqn:uw-btf} \\
	\frac{\mathrm{d} h}{\mathrm{d} x} &= - h_0 \left[ 
		\beta \cos^2 \left( \alpha x \right) \sin \left( 2 \beta x \right) +
		\alpha \cos^2 \left( \beta x \right) \sin \left( 2 \alpha x \right)
	\right]
\end{align}
Unlike the horizontal advection test, flow extends from the top of the domain all the way to the ground.  The discrete velocity field is calculated using the streamfunction in the same way as the horizontal advection test.

@@

Tracer contours at \(t = \SI{0}{\second}, \SI{5000}{\second}\) and \(\SI{10000}{\second}\) are shown in Figure~\ref{fig:advection-tracer}.
\TODO{discuss figure~\ref{fig:advection-tracer}}
Tracer errors at \(t = \SI{10000}{\second}\) are shown for the \TODO{schemes} in figures \ref{fig:advection-tracer}x and \ref{fig:advection-tracer}x respectively.

$\ell_2$ errors and tracer extrema for this test are compared with the horizontal advection results in table~\ref{tab:advection}.  In the terrain following velocity field, tracer accuracy is greatest on the BTF grid.  Errors are about ten times larger on the SLEVE and cut cell grids compared to the BTF grid.

@@

\begin{table*}
	\caption{Minimum and maximum tracer densities (\si{\kilogram\per\meter\cubed}) and \(\ell_2\) error norms, defined by equation~(\ref{eqn:l2-error}), at \(t = \SI{10000}{\second}\) in the horizontal and terrain following tracer advection tests using centred linear and cubic upwind-biased schemes.  For the horizontal advection test, \(\ell_2\) error norms, minimum and maximum values are given for the fourth order scheme using the modified code from \citet{schaer2002}.}
\label{tab:advection}
%
\centering
\footnotesize
\begin{tabular}{l l l l S S S S}
\hline\hline
                  &                     &                  & Analytic & {BTF}   & {SLEVE}         & {Cut cell}     & {No terrain} \\
\hline
Horizontal        & Centred linear      & \(\ell_2\) error & 0        & 0.284   & 0.0316          & 0.0304         & 0.0304      \\
                  &                     & min              & 0        & -0.275  & -0.0252         & -0.0251        & -0.0251     \\
                  &                     & max              & 1        & 0.925   & 0.985           & 0.985          & 0.985       \\
                  & Fourth order        & \(\ell_2\) error & 0        & 0.0938  & 0.00244         & {---}          & 0.00234     \\
                  &                     & min              & 0        & -0.0926 & -0.00174        & {---}          & -0.00178    \\
                  &                     & max              & 1        & 1.00    & 0.984           & {---}          & 0.983       \\
                  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.112   & 0.0146          & 0.00784        & 0.00784     \\
                  &                     & min              & 0        & -0.0464 & -0.0106         & -0.00674       & -0.00674    \\
                  &                     & max              & 1        & 0.922   & 0.982           & 0.983          & 0.983       \\
\hline
Terrain following & Centred linear      & \(\ell_2\) error & 0        & 0.0341  & 0.235           & 0.363          & {---}        \\
	          &                     & min              & 0        & -0.0245 & -0.120          & -1.24          & {---}        \\
		  &                     & max              & 1        & 0.985   & 0.950           & 1.01           & {---}        \\
		  & Cubic upwind-biased & \(\ell_2\) error & 0        & 0.0209  & 0.162           & 0.181          & {---}        \\
                  &                     & min              & 0        & -0.0110 & -0.0263         & -0.0284        & {---}        \\
                  &                     & max              & 1        & 0.983   & 0.865           & 0.851          & {---}        \\
\hline
\end{tabular}
\end{table*}
